---
title: "Analysis of Turgor-induced strain simulations and SOK intensity "
format: html
editor: source
editor_options: 
  chunk_output_type: console
---

# Libraries and general settings
```{r}
library(tidyverse)
library(janitor)
library(paletteer)
library(ggpubr)
library(scales)
library(kableExtra)
library(emmeans)
library(multcomp)
library(multcompView)
library(ggpmisc)

library(conflicted)
conflicts_prefer(dplyr::select, dplyr::filter, dplyr::rename)

my_palette= c("#c6c5c5", "#5fba46","#a0a0a0" )
my_palette_2 = c("#762a86", "#0481bf")
```


# Domain-specific strain analyses (Figure 2)

```{r}
simu_st5 <- read_csv("Figure 2 - Strain distribution in LRP/strain_st5.csv") |> 
  mutate(zone = fct_relevel(zone, "provascular", "boundary", "ground"))

meta_data <- simu_st5 |> 
  group_by(zone) |> 
  summarize(n=n())

Fig_Strain <- simu_st5 |> 
  ggplot(aes(y=strain*100, x=zone)) + 
  geom_boxplot(aes(fill = zone), alpha =1/3, outliers = FALSE) +
  geom_jitter(aes(color = zone), width = 0.15, size =4, alpha =2/3)+
  scale_fill_manual(values = my_palette)+
  scale_color_manual(values = my_palette)+
  scale_y_continuous(labels = label_percent(scale=1))+
  stat_compare_means(comparisons = list(c(1,2), c(2,3), c(1,3)),
                     method = "t.test",
                     label = "p.values",
                     tip.length = 0.03,
                     label.y.npc = 0.5,
                     size = 3)+
  geom_text(data = meta_data, aes(label = str_c("n=",n), y= 1))+
  theme_minimal()+
  labs(y = "Strain", 
       x= "")+
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "none")
Fig_Strain

ggsave("Fig2x.pdf", Fig_Strain, width = 8, height = 8)
```


# Strain and SOK5 signal in control and ablated conditions

## Load data

```{r}
data_ablation <- read_csv("Figure 3 - Emergence & ablations/merged.csv") |> 
  rename(edge_id = "...1") |> 
  clean_names() |> 
  mutate(stage = as.factor(stage),
         stage = fct_recode(stage, "III" = "3", "IV" = "4", "V" = "5"),
         experiment = fct_relevel(experiment, "Control"),
         edge_id = as.factor(edge_id))
```

## Metadata & stats 
```{r}
# Fit model with interaction for Strain
model_strain <- lm(strain ~ stage * experiment, data = data_ablation)
emm_strain <- emmeans(model_strain, ~ stage * experiment)

# Pairwise comparisons within each stage, Tukey-adjusted
pairs_result_strain <- pairs(emm_strain)

# Compact letter display
cld_result_strain <- multcomp::cld(emm_strain, Letters = letters, adjust = "tukey")
cld_df_strain <- as.data.frame(cld_result_strain)
```

```{r}
# Create metadata table for annotation
meta_data <- data_ablation |> 
  group_by(stage, experiment) |> 
  summarize(n=n(),
            max = max(strain),
            min= min(strain)) |> 
  left_join(cld_df_strain, by = c("stage", "experiment")) |> 
  rename(group = .group) |> 
  mutate(group = str_trim(group))
```

## Plotting

### Strain distribution for Stage 5 only (Fig 3)
```{r}
Fig_strain_st5 <- 
  data_ablation |> 
  filter(stage =="V") |> 
  ggplot(aes(y=strain, x=experiment)) + 
  geom_boxplot(aes(fill = experiment), alpha =1/3, outliers = FALSE) +
  geom_jitter(aes(color = experiment), position = position_jitterdodge(jitter.width = 0.25, dodge.width = 0.75), size =4, alpha =2/3)+
  scale_fill_manual(values = my_palette_2)+
  scale_color_manual(values = my_palette_2)+
  stat_compare_means(comparisons = list(c(1,2)),
                     method = "t.test",
                     label = "p.values",
                     tip.length = 0.03,
                     label.y.npc = 0.5,
                     size = 3)+
  geom_text(data = meta_data |> filter(stage =="V"), aes(label = str_c("n=",n), y=0.01, group = experiment))+
  theme_minimal()+ 
  labs(y = "Strain", 
       x= "")+
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "none")
Fig_strain_st5

ggsave("Fig_strain_st5.pdf", Fig_strain_st5, width = 4, height = 8)
```

### SOK5 distribution for Stage 5 only  (Fig 3)
```{r}
Fig_sok5_st5 <- 
  data_ablation |> 
  filter(stage =="V") |> 
  ggplot(aes(y=normalized_sok5, x=experiment)) + 
  geom_boxplot(aes(fill = experiment), alpha =1/3, outliers = FALSE) +
  geom_jitter(aes(color = experiment), position = position_jitterdodge(jitter.width = 0.25, dodge.width = 0.75), size =4, alpha =2/3)+
  scale_fill_manual(values = my_palette_2)+
  scale_color_manual(values = my_palette_2)+
  stat_compare_means(comparisons = list(c(1,2)),
                     method = "t.test",
                     label = "p.values",
                     tip.length = 0.03,
                     label.y.npc = 0.5,
                     size = 3)+
  geom_text(data = meta_data |> filter(stage =="V"), aes(label = str_c("n=",n), y=-0.05, group = experiment))+
  theme_minimal()+ 
  labs(y = "SOK5", 
       x= "")+
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "none")
Fig_sok5_st5

ggsave("Fig_sok5_st5.pdf", Fig_sok5_st5, width = 4, height = 8)
```


### Strain - SOK5 correlation per wall - stage 5 only (Fig S11)
```{r}
meta_data <- meta_data |> 
  mutate(y_offset = 0.18 + (as.numeric(fct_rev(experiment)) + 1) * 0.008)

Fig_strain_sok5_st5 <- 
  data_ablation |> 
  filter(stage =="V") |> 
  ggplot(aes(y=strain, x=normalized_sok5)) + 
  geom_point(aes(color = experiment), alpha =2/3, size=3) +
  geom_smooth(aes(color = experiment, group = experiment), method = "lm", se = FALSE) +
  stat_poly_eq(aes(color = experiment, group = experiment,
                    label = after_stat(paste(rr.label, p.value.label, sep = "~~~"))),
               formula = y ~ x, parse = TRUE, size = 3) +
  scale_fill_manual(values = my_palette_2)+
  scale_color_manual(values = my_palette_2)+
  geom_text(data = meta_data |> filter(stage=="V") , aes(label = str_c(experiment,": n=",n), y=y_offset, x=0.1, color = experiment, group = experiment))+
  theme_minimal()+ 
  labs(y = "Strain", 
       x= "SOK5")+
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "none")
Fig_strain_sok5_st5

ggsave("Fig_strain_sok5_st5.pdf", Fig_strain_sok5_st5, width = 8, height = 8)
```


# Strain in tissues in control vs ablated simulations (Figure S11)

## Load data (Stage 5 only)
```{r}
temp <- read_csv("Figure 3 - Emergence & ablations/st5_AB_data_set_corrected.csv") |> 
  rename(edge_id = "...1") |> 
  clean_names() |> 
  mutate(experiment = "Ablation",
         stage = "V",
         edge_id = as.factor(edge_id)) |> 
  select(edge_id, strain, normalized_strain, strain, zone, stage, experiment)

# Control conditions
temp2 <- read_csv("Figure 3 - Strain distribution in LRP/strain_st5.csv") |> 
  clean_names() |> 
  mutate(experiment = "Control", 
         stage = "V",
         edge_id = as.factor(edge_id)) |> 
  select(edge_id, strain, normalized_strain, strain, zone, stage, experiment)

data_strain_zones <- temp |> 
  rbind(temp2) |> 
  mutate(zone = fct_relevel(zone, "provascular", "boundary", "ground"),
         experiment = fct_relevel(experiment, "Control", "Ablation"))
```

## Metadata & stats 

```{r}
# Create metadata table for annotation
meta_data <- data_strain_zones |> 
  group_by(zone, experiment) |> 
  summarize(n=n(),
            max = max(strain),
            min= min(strain)) |> 
  left_join(cld_df_strain, by = c("zone", "experiment")) |> 
  rename(group = .group) |> 
  mutate(group = str_trim(group))
```

## Plotting 

### Strain distribution provs vs ground for Stage 5 only
```{r}
Fig_strain_st5_zones_byzone <- 
  data_strain_zones |> 
  ggplot(aes(y = strain, x = zone, color = zone)) + 
  geom_boxplot(aes(fill = zone), alpha = 1/3, outliers = FALSE) +
  geom_jitter(aes(color = zone), width = 0.2, size = 4, alpha = 2/3) +
  scale_fill_manual(values = my_palette) +
  scale_color_manual(values = my_palette) +
  facet_wrap(~ experiment) +
  stat_compare_means(comparisons = list(c(1, 2), c(2, 3), c(1, 3)),
                      method = "t.test",
                      label = "p.signif",
                      tip.length = 0.03,
                      label.y.npc = 0.5,
                      size = 3) +
  geom_text(data = meta_data, aes(label = str_c("n=",n), y= 0), color = "black")+
  theme_minimal() + 
  labs(y = "Strain", x = "") +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "none")
Fig_strain_st5_zones_byzone

ggsave("Fig_strain_st5_zones_byzone.pdf", Fig_strain_st5_zones_byzone, width = 8, height = 8)
```



